\chapter{Conclusion}
\label{ch:conclusion}
The first goal of this thesis is to examine the significance of secure zero-touch bootstrapping of IoT devices. We analyze and present a comparison between two bootstrapping protocols: \gls{brski} and \gls{sztp}. 
Secondly, we aim to discuss the future secrecy of cryptographic keys in IoT through analyzing the \gls{x3dh} protocol and the double ratchet algorithm. In addition to use \gls{ofmc} to formally verify a model of a \gls{x3dh} protocol variant. Moreover, provide an implementation for the future secrecy related protocols.
Thirdly, we discuss the applicability of the mentioned protocols in IoT-related use cases.
Lastly, we aimed at conducting a comparative study between certificate enrollment protocols.

At first, we presented in chapter \ref{ch:background} an overview of different concepts and cryptographic primitives, as well as related work, which are a foundation for understanding the work presented through out the thesis. Next in chapter \ref{ch:usecases}, Three use cases are introduced: \acrfull{abc}, Industrial IoT, and V2X communication. For each use case, we further discuss the relevance of secure communication and the challenges it faces. In chapter \ref{ch:secureBootstrapping}, we first give a general overview of secure bootstrapping. Afterwards, for each of the zero-touch protocols, \gls{brski} and \gls{sztp}, we give an overview of the protocol architecture, in addition to explaining the execution details for each protocol to accomplish bootstrapping of a pledge. Moreover, we compare and contrast both protocols. Furthermore, in chapter \ref{ch:postcomp}, we discussed the post-compromise security property, aka future secrecy, in the realm of secure messaging and how it differs from forward secrecy. Also, we explain two protocols, which are part of the signal protocol, that work together to achieve desirable secure messaging properties ---in the context of two party communication only--- that include future secrecy, \gls{x3dh} and the double ratchet. For the \gls{x3dh}, we additionally use \gls{ofmc} to perform a formal verification for a model of the protocol. The model is a modified version from the original specification to guarantee its security under the \gls{ofmc} intruder model. Nevertheless, our variant scarifies the deniability property offered by the original protocol. For both protocols, we discuss their post-quantum security and other worked related to their formal verification. In chapter \ref{ch:implementation}, we present a demo implementation of the protocols discussed in chapter \ref{ch:postcomp}. Chapter \ref{ch:discussion} discusses the usage of the explained protocols in chapters \ref{ch:secureBootstrapping} and \ref{ch:postcomp} in the IoT world and the use cases defined earlier. Finally, we present a comparison of certificate enrollment protocols in appendix \ref{appendix-enrollment}. 

Although the thesis does not finally conclude the discussion of IoT security, we have shown how the mentioned bootstrapping protocols achieve their goal without manual input from the user during the process. 
Despite the highlighted differences, both protocols share a large set of functionalities.
As discussed, they have differences which makes one or the other better depending on the use case. 
Regarding the X3DH protocol and the double ratchet algorithm, they are viable solutions and applicable to IoT and provide and enhance security in the proposed scenarios.
Since IoT devices are likely to be physically accessible by adversaries, the long-term keys are likely to be vulnerable. The use \gls{x3dh} protocol and the double ratchet algorithm provide longer term security against tampering of the cryptographic keys in the hardware.

