\gls{iot} is a relatively new concept that already has applications in many domains, creating new ways of interactions between humans and small devices, referred to as smart devices. Applications like Smart Airports, Transportation, Home Automation, and many more are being revolutionized by this technology \cite{marksteiner2017overview}.
\gls{iot} is the deployment of network-connected constrained devices that interact with the physical environment by employing sensors to collect data, managing other systems, controlling actuators, as well as communicating with one another. \gls{iot} devices are classified into several classes depending on their degree of computing resources and power usage. Depending on their degree, they may be assumed to not be able to process sophisticated or even conventional cryptographic operations. In many use cases, the \gls{iot} devices handle critical and private data. This imposes the requirement for data integrity and authenticity guarantees, and -because of privacy- in many cases also confidentiality.
\par
\gls{tls} \cite{rfc5246}, which relies on \gls{tcp}, has been criticized as inappropriate for constrained \gls{iot} devices \cite{shang2016challenges}. Among the reasons are, the infeasibility of \gls{iot} devices to maintain long-lived connections due to energy constraints, high header overhead, and the low-latency requirement that is opposed by the delay due to \gls{tcp} handshake, especially in a lossy network.
Alternatively, \gls{dtls} \cite{dtls} provides security for communication channels relying on \gls{udp}. In contrast to \gls{tcp}, \gls{udp} is more suitable for \gls{iot}. It is an unreliable protocol as it does not care about message delivery , resulting in a lower header overhead. In addition, it introduces less traffic to the network.
\par
The \gls{ietf} introduced \gls{coap} \cite{rfc7252} that is intended to be a generic application protocol for constrained environments. Similar to the ubiquitous HTTP \cite{http}, \gls{coap} realizes a subset of the \gls{rest} architecture. Therefore, it easily translates to HTTP. Moreover, \gls{coap} leverages \gls{udp} as its transport layer protocol which is secured by \gls{dtls}. Hence, \gls{coap} is one of well suited protocols for \gls{iot}.

signal can handle out-of-order messages (UDP) [formal analysis of signal]
